## Exercises 
The `reported_height` and `height` datasets were collected from three classes taught in the Departments of Computer Science and Biostatistics, as well as remotely through the Extension School. The biostatistics class was taught in 2016 along with an online version offered by the Extension School. On 2016-01-25 at 8:15 AM, during one of the lectures, the instructors asked students to fill in the sex and height questionnaire that populated the `reported_height` dataset. The online students filled the survey during the next few days, after the lecture was posted online. We can use this insight to define a variable, call it `type`, to denote the type of student:  `inclass` or `online`: 
```{r, eval=FALSE} 
library(lubridate) 
library(dslabs)
data("reported_heights") 
dat <- mutate(reported_heights, date_time = ymd_hms(time_stamp)) |> 
  filter(date_time >= make_date(2016, 01, 25) &  
           date_time < make_date(2016, 02, 1)) |> 
  mutate(type = ifelse(day(date_time) == 25 & hour(date_time) == 8 &  
                         between(minute(date_time), 15, 30), 
                       "inclass", "online")) |> select(sex, type) 
x <- dat$type 
y <- factor(dat$sex, c("Female", "Male")) 
``` 

1\. Show summary statistics that indicate that the `type` is predictive of sex. 

```{r}
head(dat)
table(dat)
dat %>% group_by(type, sex) %>% summarise(count = n())
# proportion female: 
26/(26+13) # 2/3 in class are female
42/(42+69) # 1/3 online are female

prop.table(table(dat))
prop.table(table(dat), 2)

# Predictive value: if type is online we would predict male and if type is inclass we would predict female. 
```

2\. Instead of using height to predict sex, use the `type` variable. 
```{r}
# Predictive value: if type is online we would predict male and if type is inclass we would predict female. 
y_hat <- ifelse(x== 'online', 'Male', 'Female')

dat$y_hat <- ifelse(x== 'online', 'Male', 'Female')
dat
```

3\. Show the confusion matrix. 
```{r}
table(y_hat, y)

table(dat$y_hat, dat$sex)
prop.table(table(dat$y_hat, dat$sex))
# Value on the diagonal are predicted corectly, values on the off diagonal are predicted wrongly.
```

4\. Use the `confusionMatrix` function in the __caret__ package to report accuracy. 
```{r}
#install.packages("caret")
library(caret)
?caret::confusionMatrix
confusionMatrix(factor(y_hat), y)

#Accuracy : 0.6333  
# check by hand:
(26+69)/(26+13+42+69)
```

5\. Now use the `sensitivity` and `specificity` functions to report specificity and sensitivity. 
```{r}
?caret::sensitivity
?caret::specificity

sensitivity(factor(y_hat), y)
specificity(factor(y_hat), y)

```

6\. What is the prevalence (% of females) in the `dat` dataset defined above? 
```{r}
mean(dat$sex == 'Female')

length(which(dat$sex == 'Female'))/nrow(dat)

# Also in the confusion atric output. 
```




