## Exercises 
1\. For each of the following, determine if the outcome is continuous or categorical: 
a. Digit reader              - categorical 
b. Movie recommendations     - continuous (if assume decimal grades are possible)
c. Spam filter               - categorical (binary: spam or not spam)
d. Hospitalizations          - continuous (not decimals, but numbers have meaning not categories)
e. Siri (speech recognition) - categorical (classify sounds into words)

2\. How many features are available to us for prediction in the digits dataset? 
```{r, eval=FALSE} 
#In the book, explains that these are 28*28 = 784 pixels. 

library(dslabs) 
mnist <- read_mnist() 
#Can also see that in mnist where data goes from 1:784

``` 

3\. In the digit reader example, the outcomes are stored here: 
```{r, eval=FALSE} 
library(dslabs) 
mnist <- read_mnist() 
y <- mnist$train$labels 
``` 

Do the following operations have a practical meaning? 
```{r, eval=FALSE} 
y[5] + y[6] 
y[5] > y[6]

# 
``` 
Pick the best answer: 
a. Yes, because $9 + 2 = 11$ and $9 > 2$. 
b. No, because `y` is not a numeric vector. 
c. No, because 11 is not a digit. It's two digits. 
d. No, because these are labels representing a category not a number. A `9` represents a class not the number 9. 
Answer: d

 
