## Exercises 
1\. Load the `GaltonFamilies` data from the __HistData__. The children in each family are listed by gender and then by height. Create a dataset called `galton_heights` by picking a male and female at random.  

```{r}
if (!require("HistData")) install.packages("HistData")
library(HistData)
library(tidyverse)
GaltonFamilies

# TA said that this question meant to ask "picking a male and female at random FROM EACH FAMILY".
galton_heights <- GaltonFamilies %>%
  group_by(family, gender) %>%
  sample_n(1)
galton_heights

```


2\. Make a scatterplot for heights between mothers and daughters, mothers and sons, fathers and daughters, and fathers and sons.

```{r}
ggplot(data = galton_heights, aes(x = mother, y = childHeight)) +
  geom_point() +
  facet_wrap(gender ~ .)

ggplot(data = galton_heights, aes(x = father, y = childHeight)) +
  geom_point() +
  facet_wrap(gender ~ .)
```

3\. Compute the correlation in heights between mothers and daughters, mothers and sons, fathers and daughters, and fathers and sons.

```{r}
#ugly way: 
#cor(galton_heights[galton_heights$gender == "female",]$father, galton_heights[galton_heights$gender == "female",]$childHeight)

galton_heights %>%
  group_by(gender) %>%
  summarise(correlation = cor(father, childHeight))

galton_heights %>%
  group_by(gender) %>%
  summarise(correlation = cor(mother, childHeight))

```


