## Exercises 
1\. Earlier we used logistic regression to predict sex from height. Use kNN to do the same. Use the code described in this chapter to select the $F_1$ measure and plot it against $k$. Compare to the $F_1$ of about 0.6 we obtained with regression. 

*Luli said this would not be on the exam in OH - someone asked what the point of this first exercise was and she said that doing logistic regression is like doing KNN with k=2 classes. if you use k=2, it sets up KNN to be similar to logistic regression output but the objectives of the two methods are different*


```{r}
library(dslabs)
library(tidyverse)
library(caret)
data("heights")

set.seed(1)
test_index <- createDataPartition(heights$sex, times = 1, p = 0.5, list = FALSE)
test_set <- heights[test_index, ]
train_set <- heights[-test_index, ]     
                
ks <- seq(1, 101, 3)
F_1 <- sapply(ks, function(k){
    fit <- knn3(sex ~ height, data = train_set, k = k)
    y_hat <- predict(fit, test_set, type = "class") %>% 
        factor(levels = levels(train_set$sex))
    F_meas(data = y_hat, reference = test_set$sex)
})
plot(ks, F_1)


#max(F_1)
```




2\. Load the following dataset: 
```{r, eval=FALSE} 
data("tissue_gene_expression") 
``` 
This dataset includes a matrix `x`:  
```{r, eval=FALSE} 
dim(tissue_gene_expression$x) 

#189 samples and 500 genes
``` 
with the gene expression measured on 500 genes for 189 biological samples representing seven different tissues. The tissue type is stored in `y`: 
```{r, eval=FALSE} 
table(tissue_gene_expression$y) 
``` 
Split the data in training and test sets, then use kNN to predict tissue type and see what accuracy you obtain. Try it for  $k = 1, 3, \dots, 11$. 

Q from OH: How do we interpret the results of KNN? Is this accuracy?
Luli: Yes this tells us the accuracy and that no matter which k you pick it's pretty accurate, since they are all close to 1

```{r}

# From OH - Luli said she would want to take into account trying to get a similar amount from each group, but that this isn't necessary for this - can just get a random 80-20 split
#?knn3
#length(tissue_gene_expression$y)   length is 189
# taking sample from 1:length(tissue_gene_expression$y) to creating a 80-20 split; usually do 80% train, 20% test 
train_indx <-sample(c(1:189), round(0.8*189))
test_indx <-c(1:189)[-train_indx]
  
for (k in seq(1,11,2)){
  fit <- knn3(tissue_gene_expression$x[train_indx,], 
              tissue_gene_expression$y[train_indx], 
              k=k)
  
  y_hat<-predict(fit,
                 tissue_gene_expression$x[test_indx,],
                 type = "class") # always fitted model and then the new data - just need to input X's for the prediction
  print(paste0('k=',k))
  print(mean(tissue_gene_expression$y[test_indx]==y_hat))
}

## ALTERNATIVE METHOD
set.seed(1)
library(caret)
y <- tissue_gene_expression$y
x <- tissue_gene_expression$x
train_index <- createDataPartition(y, list = FALSE)

train_index
sapply(seq(1, 11, 2), function(k){
    fit <- knn3(x[train_index,], y[train_index], k = k)
    y_hat <- predict(fit, newdata = data.frame(x=x[-train_index,]),
                type = "class") ## specifting in the test data
mean(y_hat == y[-train_index])
})
```


*OH QUESTIONS*
Question during OH: For predict knn it can't give you the class but for predict linear or logistic it can only give you the probability, how do we decide the division rule is it always 0.5?

Response:  Linear regression uses RMSE which is continuous metric. For logistic regression, set a cutoff if you're interested in comparing the classifications, also it's arbitrary, depending on what you think is fair. Most people use 0.5 by default; if greater than 0.5 you have a higher chance of being one class than the other. During the exam, then the cutoff would be specified

Question: will we run code during the exam? 
Response: Some questions provide the code, have us run it and then look at teh output and answer the question
Might have to take an average or standard deviation, but won't have to write a lot of code

Question: will it be like the last exam?
Many will be very similar. For string processing, if there is a questino it will be a simpler here's an expression, which pattern matches this regular expresses, rather than changing a bunch of data

For the exam, look at output, from the output what does it say the accuracy is

 
