## Exercises 
1\. In the wrangling part of this book, we used the code below to obtain mortality counts for Puerto Rico for 2015-2018. 
```{r, eval=FALSE} 
library(tidyverse) 
library(lubridate) 
library(purrr) 
library(pdftools) 
library(dslabs) 
fn <- system.file("extdata", "RD-Mortality-Report_2015-18-180531.pdf", 
                  package="dslabs") 
dat <- map_df(str_split(pdf_text(fn), "\n"), function(s){ 
  s <- str_trim(s) 
  header_index <- str_which(s, "2015")[1] 
  tmp <- str_split(s[header_index], "\\s+", simplify = TRUE) 
  month <- tmp[1] 
  header <- tmp[-1] 
  tail_index  <- str_which(s, "Total") 
  n <- str_count(s, "\\d+") 
  out <- c(1:header_index, which(n == 1),  
           which(n >= 28), tail_index:length(s)) 
  s[-out] %>%  str_remove_all("[^\\d\\s]") %>% str_trim() %>% 
    str_split_fixed("\\s+", n = 6) %>% .[,1:5] %>% as_tibble() %>%  
    setNames(c("day", header)) %>% 
    mutate(month = month, day = as.numeric(day)) %>% 
    pivot_longer(-c(day, month), names_to = "year", values_to = "deaths") %>% 
    mutate(deaths = as.numeric(deaths)) 
}) %>% 
  mutate(month = recode(month,  
                        "JAN" = 1, "FEB" = 2, "MAR" = 3,  
                        "APR" = 4, "MAY" = 5, "JUN" = 6,  
                        "JUL" = 7, "AGO" = 8, "SEP" = 9,  
                        "OCT" = 10, "NOV" = 11, "DEC" = 12)) %>% 
  mutate(date = make_date(year, month, day)) %>% 
  filter(date <= "2018-05-01") 
``` 
Use the `loess` function to obtain a smooth estimate of the expected number of deaths as a function of date. Plot this resulting smooth function. Make the span about two months long. 

```{r}
dat
?loess
any(is.na(dat$death))

myloess <- loess(deaths ~ as.numeric(date), dat, span = 60/as.numeric(diff(range(dat$date))))
myloess
#plot(myloess)

dat %>% mutate(prediction = predict(myloess, as.numeric(date))) %>%
  ggplot(aes(x = date)) +
  geom_point(aes(y = deaths)) + 
  geom_line(aes(y = prediction), color = "blue") # bc blue is mai-han's fave color

dat %>% 
  ggplot(aes(x = date, y = deaths)) +
  geom_point() + 
  geom_smooth(method='loess', span = 60/as.numeric(diff(range(dat$date))))

```

2\. Plot the smooth estimates against day of the year, all on the same plot but with different colors. 
```{r}
# dat %>% mutate(prediction = predict(myloess, as.numeric(date)),
#                daymonth = as.Date(date, "%m-%d")) %>%
#   ggplot() +
#   geom_point(aes(x = daymonth, y = deaths)) + 
#   geom_line(aes(x = daymonth, y = prediction, color = year)) 

dat %>% mutate(prediction = predict(myloess, as.numeric(date)),
                       daymonth2 = as.numeric(date) - (as.numeric(year)-2015)*365 ) %>% 
  ggplot() + 
  geom_line(aes(x = daymonth2, y = prediction, color = year))

```


3\. Suppose we want to predict 2s and 7s in our `mnist_27` dataset with just the second covariate. Can we do this? On first inspection it appears the data does not have much predictive power. In fact, if we fit a regular logistic regression, the coefficient for `x_2` is not significant! 
```{r, eval = FALSE} 
library(broom) 
library(dslabs) 
data("mnist_27") 
mnist_27$train |>  
  glm(y ~ x_2, family = "binomial", data = _) |>  
  tidy() 
``` 
Plotting a scatterplot here is not useful since `y` is binary: 
```{r, eval = FALSE} 
qplot(x_2, y, data = mnist_27$train) 
``` 
Fit a loess line to the data above and plot the results. Notice that there is predictive power, except the conditional probability is not linear. 

```{r}
head(mnist_27$train)

ggplot(data = mnist_27$train,aes(x = x_2, y = y)) +
  geom_point( ) +
  geom_smooth(method='loess')

# I'm not sure about this answer. 

#maddy attempt
my_loess2<-loess(as.numeric(y) ~ as.numeric(x_2), mnist_27$train)#default span is 0.75
mnist_27$test |> 
  mutate(prediction = predict(my_loess2, as.numeric(x_2))) |>
  ggplot(aes(x=x_2)) + 
  geom_point(aes(y = y)) + 
  geom_line(aes(y = prediction), color = 'red')

```

