
## Exercises

1\. Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N=25$. Consider the random variable $S$ which is the **total** number of Democrats in your sample. What is the expected value of this random variable? Hint: it's a function of $p$.

$$
S = N*p
$$

2\. What is the standard error of $S$ ? Hint: it's a function of $p$.

$$
\sqrt{N*p*(1-p)}
$$

3\. Consider the random variable $S/N$. This is equivalent to the sample average, which we have been denoting as $\bar{X}$. What is the expected value of the $\bar{X}$? Hint: it's a function of $p$.

$$
p
$$

4\. What is the standard error of $\bar{X}$? Hint: it's a function of $p$.

$$
\sqrt{p*(1-p)/N}
$$

5\. Write a line of code that gives you the standard error `se` for the problem above for several values of $p$, specifically for `p <- seq(0, 1, length = 100)`. Make a plot of `se` versus `p`.

```{r}
p <- seq(0, 1, length = 100)
p
N <- 25
se_p <- sqrt(p*(1-p) / N)
se_p

library(ggplot2)
dat <- data.frame(p = p, sep_p = se_p)
ggplot(dat) + 
  geom_point(aes(x = se_p, y = p)) + 
  theme_bw()

```

6\. Copy the code above and put it inside a for-loop to make the plot for $N=25$, $N=100$, and $N=1000$.

```{r}
library(gridExtra)
N <- c(25, 100, 1000)
g <- NULL

for (i in 1:length(N)) {
  p <- seq(0, 1, length = 100)
  se_p <- sqrt(p*(1-p) / N[i])
  dat <- data.frame(p = p, sep_p = se_p)
  
  g[[i]] <- 
    ggplot(dat, aes(x = sep_p, y = p)) + 
    geom_point() + 
    theme_bw()
}

grid.arrange(grobs = g, ncol= 3)

# X axis is different on plots !! 
# As larger N, standard error is smaller. 

```

```{r}
library(gridExtra)
N <- c(25, 100, 1000)
dat <- NULL

for (i in 1:length(N)) {
  p <- seq(0, 1, length = 100)
  se_p <- sqrt(p*(1-p) / N[i])
  tmp <- data.frame(p = p, sep_p = se_p, N = N[i])
  dat <- rbind(dat, tmp)
}

ggplot(dat, aes(x = sep_p, y = p)) +
  geom_point() + 
  facet_wrap(~N) +
  theme_bw()

# As larger N, standard error is smaller. 

```

7\. If we are interested in the difference in proportions, $p - (1-p)$, our estimate is $d = \bar{X} - (1-\bar{X})$. Use the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$.

$$
p - (1-p) = 2p - 1

$$

8\. What is the standard error of $d$?

$$

\sqrt{p(1-p)/N + (1-p)(1-(1-p))/N} 
= \sqrt{[p(1-p) + p(1-p)]/N}

#maihan:
SE[d]=SE[ \bar{X} - (1-\bar{X})] = SE[2\bar{X}]-SE[1]=2SE[\bar{X}]=2sqrt(p(1-p)/N)

$$
*maddy addition*
This should simplify to $$\sqrt{2*p(1-p)/N}$$


9\. If the actual $p=.45$, it means the Republicans are winning by a relatively large margin since $d= -.1$, which is a 10% margin of victory. In this case, what is the standard error of $2\hat{X}-1$ if we take a sample of $N=25$?

```{r}
N <- 25
p <- 0.45
sqrt(p*(1-p) + p*(1-p))/N

#maddy comment: I think the N should be included in the square root here!
#if we recalculate the SE with square rooted N:
sqrt((p*(1-p) + p*(1-p))/N)
#now the SE is 0.14, which is larger than d
```

10\. Given the answer to 9, which of the following best describes your strategy of using a sample size of $N=25$?

a.  The expected value of our estimate $2\bar{X}-1$ is $d$, so our prediction will be right on.
b.  Our standard error is larger than the difference, so the chances of $2\bar{X}-1$ being positive and throwing us off were not that small. We should pick a larger sample size.
c.  The difference is 10% and the standard error is about 0.2, therefore much smaller than the difference.
d.  Because we don't know $p$, we have no way of knowing that making $N$ larger would actually improve our standard error.

```{r}
# c ? 
#Maddy: per my answer above, I think b is correct!

```